HO unification from object language to meta language

### 2024-03-21 paper/appendix.tex
build

\begin{elpicode}
  type use-binder fo.fm -> fo.fm -> o.
  use-binder N N.
  use-binder N (fo.fapp L) :- exists (use-binder N) L.
  use-binder N (fo.flam B) :- pi x\ use-binder N (B x).

  type maybe-eta fo.fm -> fo.fm -> list fo.fm -> o.
  maybe-eta N (fo.fapp[fo.fuva _|Args]) _ :- !,
    exists (x\ maybe-eta-of [] N x) Args, !. 
  maybe-eta N (fo.flam B) L :- !, pi x\ maybe-eta N (B x) [x | L].
  maybe-eta _ (fo.fapp [fo.fcon _|Args]) L :-
    split-last-n {len L} Args First Last,
    forall1 (x\ forall1 (y\ not (use-binder x y)) First) L,
    forall2 (maybe-eta-of []) {std.rev L} Last.

  type maybe-eta-of list fo.fm -> fo.fm -> fo.fm -> o.
  maybe-eta-of _ N N :- !.
  maybe-eta-of L N (fo.fapp[fo.fuva _|Args]) :- !, 
    % Head is flex -> if N is in Args, we can have eta
    forall1 (x\ exists (maybe-eta-of [] x) Args) [N|L]. 
  maybe-eta-of L N (fo.flam B) :- !, 
    pi x\ maybe-eta-of [x | L] N (B x).
  maybe-eta-of L N (fo.fapp [N|Args]) :-
    % Head is rigid -> to be an eta we consider also L
    last-n {len L} Args R,
    forall2 (maybe-eta-of []) R {std.rev L}.

  type locally-bound ho.tm -> o.
  type get-scope-aux ho.tm -> list ho.tm -> o.
  get-scope-aux (ho.con _) [].
  get-scope-aux (ho.uva _ L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux (ho.lam B) L1 :- 
    pi x\ locally-bound x => get-scope-aux (B x) L1.
  get-scope-aux (ho.app L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux X [X] :- name X, not (locally-bound X).
  get-scope-aux X [] :- name X, (locally-bound X).

  type get-scope ho.tm -> list ho.tm -> o.
  get-scope T Scope :- names N,
    get-scope-aux T ScopeDuplicata,
    std.filter N (mem ScopeDuplicata) Scope.

  type close-links (ho.tm -> links) -> links -> o.
  close-links (_\[]) [].
  close-links (v\[L|XS v]) [L|YS] :- !, close-links XS YS.
  close-links (v\[(L v)|XS v]) [ho.abs L|YS] :- !, 
    close-links XS YS.

  type comp-lam (fo.fm -> fo.fm) -> (ho.tm -> ho.tm) -> 
    mappings -> mappings -> links -> links -> ho.subst -> 
      ho.subst -> o.
  comp-lam F F1 Mappings1 Mappings2 L L2 S S1 :-
    pi x y\ (pi M L S\ comp x y M M L L S S) => 
      comp (F x) (F1 y) Mappings1 Mappings2 L (L1 y) S S1, 
        close-links L1 L2.

  type comp fo.fm -> ho.tm -> mappings -> mappings -> links -> 
    links -> ho.subst -> ho.subst -> o.
  comp (fo.fcon X) (ho.con X)        Map1 Map1 L1 L1 S1 S1.
  comp (fo.flam F) (ho.uva E0 Scope) Map1 Map2 L1 L3 S1 S3 :-
    (pi x\ maybe-eta x (F x) [x]), !,
      mem.new S1 E0 S2,
      comp-lam F F1 Map1 Map2 L1 L2 S2 S3,
      (pi x\ locally-bound x => get-scope (F1 x) Scope), 
      NewLink = ho.val(link-eta (ho.uva E0 Scope) (ho.lam F1)), 
      L3 = [NewLink| L2].
  comp (fo.flam F) (ho.lam F1)       Map1 Map2 L1 L2 S1 S2 :- 
    comp-lam F F1 Map1 Map2 L1 L2 S1 S2.
  comp (fo.fuva N) (ho.uva M []) Map1 Map2 L L S S1 :- 
    alloc Map1 Map2 (fv N) (hv M (arity z)) S S1.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M Scope) Map1 Map2 L1 L1 S1 S2 :-
    pattern-fragment Ag, !,
      fold6 comp Ag Scope Map1 Map1 L1 L1 S1 S1,
      len Ag Arity, 
      alloc Map1 Map2 (fv N) (hv M (arity Arity)) S1 S2.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M' Scope) Map1 Map3 L1 L3 S1 S4 :- !,
    pattern-fragment-prefix Ag PF NPF,
      fold6 comp PF PF1 Map1 Map1 L1 L1 S1 S1,
      fold6 comp NPF NPF1 Map1 Map2 L1 L2 S1 S2,
      len PF Arity, 
      alloc Map2 Map3 (fv N) (hv M (arity Arity)) S2 S3,
      get-scope (ho.app {append PF1 NPF1}) Scope, 
      mem.new S3 M' S4,
      L3 = [@val-link-beta (ho.uva M' Scope) (ho.app[ho.uva M PF1 | NPF1]) | L2].
  comp (fo.fapp A) (ho.app A1) Map1 Map2 L1 L2 S1 S2 :- 
    fold6 comp A A1 Map1 Map2 L1 L2 S1 S2.

  type make-eta-link-aux nat -> address -> address -> 
    list ho.tm -> links -> ho.subst -> ho.subst -> o.
  make-eta-link-aux z Ad1 Ad2 Scope1 L H1 H1  :-
    std.rev Scope1 Scope, eta-expand (ho.uva Ad2 Scope) @one T1,
    L = [@val-link-eta (ho.uva Ad1 Scope) T1].
  make-eta-link-aux (s N) Ad1 Ad2 Scope1 L H1 H3 :-
    std.rev Scope1 Scope, mem.new H1 Ad H2, 
    eta-expand (ho.uva Ad Scope) @one T2,
    (pi x\ make-eta-link-aux N Ad Ad2 [x|Scope1] (L1 x) H2 H3),
    close-links L1 L2,
    L = [@val-link-eta (ho.uva Ad1 Scope) T2 | L2].

  type make-eta-link nat -> nat -> address -> address -> 
          list ho.tm -> links -> ho.subst -> ho.subst -> o.
  make-eta-link (s N) z Ad1 Ad2 Vars L H H1 :- 
    make-eta-link-aux N Ad2 Ad1 Vars L H H1.
  make-eta-link z (s N) Ad1 Ad2 Vars L H H1 :- 
    make-eta-link-aux N Ad1 Ad2 Vars L H H1.
  make-eta-link (s N) (s M) Ad1 Ad2 Vars Links H H1 :-
    (pi x\ make-eta-link N M Ad1 Ad2 [x|Vars] (L x) H H1),
    close-links L Links.

  type deduplicate-mappings mappings -> mappings -> 
          ho.subst -> ho.subst -> links -> links -> o.
  deduplicate-mappings [] [] H H L L.
  deduplicate-mappings [(mapping (fv O) (hv M (arity LenM)) as X1) | Map1] Map2 H1 H3 L1 L3 :-
    take-list Map1 (mapping (fv O) (hv M' (arity LenM'))) _, !,
    make-eta-link LenM LenM' M M' [] New H1 H2,
    append New L1 L2,
    deduplicate-mappings Map1 Map2 H2 H3 L2 L3.
  deduplicate-mappings [A|As] [A|Bs] H1 H2 L1 L2 :- 
    deduplicate-mappings As Bs H1 H2 L1 L2, !.
\end{elpicode}

>>>>

\input{code/comp}
